---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r import-data, cache=TRUE}
actdata<-read.table(unz("activity.zip","activity.csv"),header=T,sep=",")
```

## What is mean total number of steps taken per day?
1.1 - Calculating the total number of steps per day and graphing a histogram with the data  

```{r cache=TRUE}
totsteps<-aggregate(steps~date,data=actdata,FUN=sum,na.rm=TRUE)
hist(totsteps$steps,breaks = 20)  
meantotsteps<-mean(totsteps$steps)
mediantotsteps<-median(totsteps$steps)

```
1.2 - Calculating the mean steps per day

Mean of total number of steps is `r meantotsteps`

1.3 - Calculating the mean steps per day

Median of total number of steps is `r mediantotsteps`


## What is the average daily activity pattern?

2.1 - Calculating average by intervals and plotting it
```{r cache=TRUE}
intervalmean<-aggregate(steps~interval,data=actdata,FUN=mean,na.rm=TRUE)
plot(intervalmean$interval,intervalmean$steps,type="l")
intervalmaxmean<-intervalmean[which.max(intervalmean$steps),]$interval
```

2.2 - The interval with maximum average number of steps is `r intervalmaxmean`

## Imputing missing values

```{r}
missingval<-sum(is.na(actdata$steps))
```

3.1 - Number of mising values in the dataset are `r missingval`

3.2 - 
Using a strategy for filing in missing values with the mean of intervals. 

Adding a new function to return mean steps for a specific interval
```{r}
meanstepsbyint<-function(interval){
    intervalmean[intervalmean$interval==interval,]$steps
}
```

3.3 - 
Copying dataset to a new dataset
```{r}
actdata2<-actdata
```

Replacing NAs in new dataset
```{r}
counta=0
for(i in 1:nrow(actdata2)){
  if(is.na(actdata2[i,]$steps)){
    actdata2[i,]$steps<-meanstepsbyint(actdata2[i,]$interval)
    counta=counta+1
  }
}
```

`r counta` values replaced with mean steps by interval

3.4 - Calculating the total number of steps per day and graphing a histogram with the new dataset with no missing values  

```{r cache=TRUE}
totsteps2<-aggregate(steps~date,data=actdata2,FUN=sum,na.rm=TRUE)
hist(totsteps2$steps,breaks = 20)  
meantotsteps2<-mean(totsteps2$steps)
mediantotsteps2<-median(totsteps2$steps)

```

The mean total steps per day with new dataset is `r meantotsteps2`

The median total steps per day with new dataset is `r mediantotsteps2`

There mean before and after imputing data stays same with a very little difference with median



## Are there differences in activity patterns between weekdays and weekends?

Converting date field to type date from a character
```{r cache=TRUE}
actdata2$date<-as.Date(actdata2$date)
```

4.1 - 
Adding a field to identify day of the week and then categorize it into weekday or weekend
```{r cache=TRUE}
actdata2$day<-weekdays(actdata2$date)
for(i in 1:nrow(actdata2)){
    if(actdata2[i,]$day=="Saturday"||actdata2[i,]$day=="Sunday"){
        actdata2[i,]$day="Weekend"
    }
    else{
        actdata2[i,]$day="Weekday"
    }
}
```


4.2 - 
Creating a mean of steps by interval and weekday/weekend
```{r cache=TRUE}
intervalwdaymean<-aggregate(steps~interval+day,data=actdata2,FUN=mean,na.rm=TRUE)
```

4.3 - Panel plot
```{r cache=TRUE}
library(lattice)
xyplot(steps~interval|factor(day),data=actdata2,aspect=1/2,type="l")
```

